Modal Abstractions for Virtualizing Memory Addresses

### 2023-11-16 examples.tex
fix build failures

\begin{figure}\footnotesize
\begin{lstlisting}[mathescape]
$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ulcorner \textsf{KERNELBASE} + \paddr > 0 \urcorner \ast \texttt{rbx}\mapsto_{\textsf{r}} \paddr \ast \sumapaces\rtv\delta  \ast \sumpv\paddr\qfrac\true  }_{\rtv}$
add rbx, KERNELBASE ;; shift physical address by identity mapping constant
$\specline{\exists \qfraczero; \qfrac,\textsf{ w, val}, \textsf{P} \ast \texttt{rbx}\mapsto_{\textsf{r}} \paddr + \textsf{KERNELBASE} }_{\rtv}$
$\specline{\bigast{(\vaddr, \paddr)\in \ptablestore}{\exists\;(\textsf{l4e, l3e, l2e, l1e, } \paddr)\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr\textsf{, l4e, l3e, l2e, l1e,} \paddr)}}_{\rtv}$
$\specline{\mathcal{R} -\ast \bigast{(\paddr) \in \Xi}{\exists\; (\textsf{qfrac, q, val, va}) \ldotp \ulcorner \textsf{va} = \paddr + \textsf{KERNELBASE} \urcorner \ast \sumwalkabs\vaddr\qfrac\paddr  \ast \paddr \mapsto_{\mathsf{p}}\{\textsf{qfrac}\} \; \vale\ast  \ulcorner \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }(\vale) \urcorner}}_{\rtv}$
$\specline{ \ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}} \ast \ownGhostpv\gammaPred{\authfull{\pvmapping\Xi}}}$
$\specline{ \sumapaces\rtv\delta \ast \sumpv\paddr\qfraczero\true \ast \textsf{P}}_{\rtv}$
$\specline{\vaddr \mapsto_{\texttt{vpte}} \{\textsf{q}, \textsf{qfrac}\} \paddr \;(\texttt{val})}$
\end{lstlisting}
\todo[inline]{Ismail: Should \delta be something else? $\delta$ was used for $\theta$, we need a different gname for $\Xi$.}
\caption{Converting a physical address of a PTE to a virtual address (without instruction pointer or flag updates).}
\label{fig:p2v}
\end{figure}

@@

\subsection{Change of Address Space}
A critical piece of \emph{trusted} code in current verified OS kernels is the assembly code to change the current address space; current verified OS kernels currently 
lack effective ways to specify and reason about this low-level operation, for reasons outlined in Section \ref{sec:relwork}.
\begin{figure}\footnotesize
\begin{lstlisting}[mathescape]
;; Assume the save-space is in rdi, load-space in rsi. First, save the yielding context
$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast [\rtv'](\mathcal{I}\texttt{ASpace}(\theta',m') \ast \texttt{Pother}) \ast \texttt{rsi}\mapsto_{\textsf{r}} old \ast \texttt{rdi}\mapsto_{\tnewoldr}} old \ast \texttt{rbx}\mapsto_{\textsf{r}} \texttt{rbxv} }_{\rtv}$
$\specline{\texttt{rsp}\mapsto_{\textsf{r}} \texttt{rspv} \ast \texttt{rbp}\mapsto_{\textsf{r}} \texttt{rbpv} \ast \texttt{r12}\mapsto_{\textsf{r}} \texttt{r12v} \ast \texttt{r13}\mapsto_{\textsf{r}} \texttt{r13v} \ast \texttt{r14}\mapsto_{\textsf{r}} \texttt{r14v} \ast \texttt{r15}\mapsto_{\textsf{r}} \texttt{r15v}}_{\rtv}$
$\specline{\mathsf{ContextAt}(old,\_) \ast \mathsf{ContextAt}(new,[\mathsf{rbxv}',\ldots,\rtv']) }_{\rtv}$
mov 0[rdi], rbx
... ;; mov rsp, rbp, r12, r13, r14, saved to offsets 8, 16, 24, 32, and 40 from rdi
mov 48[rdi], r15
$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast [\rtv'](\mathcal{I}\texttt{ASpace}(\theta',m') \ast \texttt{Pother})  \ast  \texttt{rsi}\mapsto_{\textsf{r}} new \ast \texttt{rdi}\mapsto_{\textsf{r}} old \ast \texttt{rbx}\mapsto_{\textsf{r}} \texttt{rbxv} }_{\rtv}$
$\specline{ \texttt{rsp}\mapsto_{\textsf{r}} \texttt{rspv} \ast \texttt{rbp}\mapsto_{\textsf{r}} \texttt{rbpv} \ast \texttt{r12}\mapsto_{\textsf{r}} \texttt{r12v} \ast \texttt{r13}\mapsto_{\textsf{r}} \texttt{r13v} \ast \texttt{r14}\mapsto_{\textsf{r}} \texttt{r14v} \ast \texttt{r15}\mapsto_{\textsf{r}} \texttt{r15v}}_{\rtv}$
$\specline{\mathsf{ContextAt}(old,[\mathsf{rbxv,\ldots,\rtv}]) \ast \mathsf{ContextAt}(new,[\mathsf{rbxv}',\ldots,\rtv']) }_{\rtv}$
mov 56[%rdi], %cr3
$\specline{\ldots \ast \texttt{rdi+56} \mapsto_{\textsf{v}} \rtv}_{\rtv}$    
;; Restore target context
mov rbx, 0[rsi] 
mov rsp, 8[rsi] ;; Switch to new stack, which may not be mapped in the current address space!
... ;; load rbp, r12, r13, r14, from offsets 16, 24, 32, and 40 from rsi
mov r15, 48[rsi]
$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast [\rtv'](\mathcal{I}\texttt{ASpace}(\theta',m') \ast \texttt{Pother})  \ast  \texttt{rsi}\mapsto_{\textsf{r}} new \ast \texttt{rdi}\mapsto_{\textsf{r}} old \ast \texttt{rbx}\mapsto_{\textsf{r}} \texttt{rbxv}' }_{\rtv}$
$\specline{\texttt{rsp}\mapsto_{\textsf{r}} \texttt{rspv}' \ast \texttt{rbp}\mapsto_{\textsf{r}} \texttt{rbpv}' \ast \texttt{r12}\mapsto_{\textsf{r}} \texttt{r12v}' \ast \texttt{r13}\mapsto_{\textsf{r}} \texttt{r13v}' \ast \texttt{r14}\mapsto_{\textsf{r}} \texttt{r14v}' \ast \texttt{r15}\mapsto_{\textsf{r}} \texttt{r15v}'}_{\rtv}$
$\specline{\mathsf{ContextAt}(old,[\mathsf{rbxv,\ldots,\rtv}]) \ast \mathsf{ContextAt}(new,[\mathsf{rbxv}',\ldots,\rtv']) }_{\rtv}$
;; Switch to the new address space
mov cr3, 56[rsi]
$\specline{ [\rtv](\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast \mathsf{ContextAt}(old,[\mathsf{rbxv,\ldots,\rtv}]) \ast \mathsf{ContextAt}(new,[\mathsf{rbxv}',\ldots,\rtv']))  }_{\rtv}$
$\specline{ \mathcal{I}\texttt{ASpace}(\theta',m') \ast \texttt{Pother} \ast  \texttt{rsi}\mapsto_{\textsf{r}} new \ast \texttt{rdi}\mapsto_{\textsf{r}} old \ast \ldots }_{\rtv'}$
\end{lstlisting}
\vspace{-1em}
\caption{Basic task switch code that switches address spaces.}
\label{fig:swtch}
\end{figure}

>>>>

\begin{figure}\footnotesize
\begin{lstlisting}[mathescape]
$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ulcorner \textsf{KERNELBASE} + \paddr > 0 \urcorner \ast \texttt{rbx}\mapsto_{\textsf{r}} \paddr \ast \sumapaces\rtv\delta  \ast \sumpv\paddr\qfrac\true  }_{\rtv}$
add rbx, KERNELBASE ;; shift physical address by identity mapping constant
$\specline{\exists \qfraczero; \qfrac,\textsf{ w, val}, \textsf{P} \ast \texttt{rbx}\mapsto_{\textsf{r}} \paddr + \textsf{KERNELBASE} }_{\rtv}$
$\specline{\bigast{(\vaddr, \paddr)\in \ptablestore}{\exists\;(\textsf{l4e, l3e, l2e, l1e, } \paddr)\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr\textsf{, l4e, l3e, l2e, l1e,} \paddr)}}_{\rtv}$
$\specline{\mathcal{R} -\ast \bigast{(\paddr) \in \Xi}{\exists\; (\textsf{qfrac, q, val, va}) \ldotp \ulcorner \textsf{va} = \paddr + \textsf{KERNELBASE} \urcorner \ast \sumwalkabs\vaddr\qfrac\paddr  \ast \paddr \mapsto_{\mathsf{p}}\{\textsf{qfrac}\} \; \vale\ast  \ulcorner \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }(\vale) \urcorner}}_{\rtv}$
$\specline{ \ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}} \ast \ownGhostpv\gammaPred{\authfull{\pvmapping\Xi}}}$
$\specline{ \sumapaces\rtv\delta \ast \sumpv\paddr\qfraczero\true \ast \textsf{P}}_{\rtv}$
$\specline{\vaddr \mapsto_{\texttt{vpte}} \{\textsf{q}, \textsf{qfrac}\} \paddr \;(\texttt{val})}$
\end{lstlisting}
\todo[inline]{Ismail: Should $\delta$ be something else? $\delta$ was used for $\theta$, we need a different gname for $\Xi$.}
\caption{Converting a physical address of a PTE to a virtual address (without instruction pointer or flag updates).}
\label{fig:p2v}
\end{figure}

@@

\subsection{Change of Address Space}
A critical piece of \emph{trusted} code in current verified OS kernels is the assembly code to change the current address space; current verified OS kernels currently 
lack effective ways to specify and reason about this low-level operation, for reasons outlined in Section \ref{sec:relwork}.
\begin{figure}\footnotesize
\begin{lstlisting}[mathescape]
;; Assume the save-space is in rdi, load-space in rsi. First, save the yielding context
$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast [\rtv'](\mathcal{I}\texttt{ASpace}(\theta',m') \ast \texttt{Pother}) \ast \texttt{rsi}\mapsto_{\textsf{r}} new \ast \texttt{rdi}\mapsto_{\textsf{r}} old \ast \texttt{rbx}\mapsto_{\textsf{r}} \texttt{rbxv} }_{\rtv}$
$\specline{\texttt{rsp}\mapsto_{\textsf{r}} \texttt{rspv} \ast \texttt{rbp}\mapsto_{\textsf{r}} \texttt{rbpv} \ast \texttt{r12}\mapsto_{\textsf{r}} \texttt{r12v} \ast \texttt{r13}\mapsto_{\textsf{r}} \texttt{r13v} \ast \texttt{r14}\mapsto_{\textsf{r}} \texttt{r14v} \ast \texttt{r15}\mapsto_{\textsf{r}} \texttt{r15v}}_{\rtv}$
$\specline{\mathsf{ContextAt}(old,\_) \ast \mathsf{ContextAt}(new,[\mathsf{rbxv}',\ldots,\rtv']) }_{\rtv}$
mov 0[rdi], rbx
... ;; mov rsp, rbp, r12, r13, r14, saved to offsets 8, 16, 24, 32, and 40 from rdi
mov 48[rdi], r15
$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast [\rtv'](\mathcal{I}\texttt{ASpace}(\theta',m') \ast \texttt{Pother})  \ast  \texttt{rsi}\mapsto_{\textsf{r}} new \ast \texttt{rdi}\mapsto_{\textsf{r}} old \ast \texttt{rbx}\mapsto_{\textsf{r}} \texttt{rbxv} }_{\rtv}$
$\specline{ \texttt{rsp}\mapsto_{\textsf{r}} \texttt{rspv} \ast \texttt{rbp}\mapsto_{\textsf{r}} \texttt{rbpv} \ast \texttt{r12}\mapsto_{\textsf{r}} \texttt{r12v} \ast \texttt{r13}\mapsto_{\textsf{r}} \texttt{r13v} \ast \texttt{r14}\mapsto_{\textsf{r}} \texttt{r14v} \ast \texttt{r15}\mapsto_{\textsf{r}} \texttt{r15v}}_{\rtv}$
$\specline{\mathsf{ContextAt}(old,[\mathsf{rbxv,\ldots,\rtv}]) \ast \mathsf{ContextAt}(new,[\mathsf{rbxv}',\ldots,\rtv']) }_{\rtv}$
mov 56[%rdi], %cr3
$\specline{\ldots \ast \texttt{rdi+56} \mapsto_{\textsf{v}} \rtv}_{\rtv}$    
;; Restore target context
mov rbx, 0[rsi] 
mov rsp, 8[rsi] ;; Switch to new stack, which may not be mapped in the current address space!
... ;; load rbp, r12, r13, r14, from offsets 16, 24, 32, and 40 from rsi
mov r15, 48[rsi]
$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast [\rtv'](\mathcal{I}\texttt{ASpace}(\theta',m') \ast \texttt{Pother})  \ast  \texttt{rsi}\mapsto_{\textsf{r}} new \ast \texttt{rdi}\mapsto_{\textsf{r}} old \ast \texttt{rbx}\mapsto_{\textsf{r}} \texttt{rbxv}' }_{\rtv}$
$\specline{\texttt{rsp}\mapsto_{\textsf{r}} \texttt{rspv}' \ast \texttt{rbp}\mapsto_{\textsf{r}} \texttt{rbpv}' \ast \texttt{r12}\mapsto_{\textsf{r}} \texttt{r12v}' \ast \texttt{r13}\mapsto_{\textsf{r}} \texttt{r13v}' \ast \texttt{r14}\mapsto_{\textsf{r}} \texttt{r14v}' \ast \texttt{r15}\mapsto_{\textsf{r}} \texttt{r15v}'}_{\rtv}$
$\specline{\mathsf{ContextAt}(old,[\mathsf{rbxv,\ldots,\rtv}]) \ast \mathsf{ContextAt}(new,[\mathsf{rbxv}',\ldots,\rtv']) }_{\rtv}$
;; Switch to the new address space
mov cr3, 56[rsi]
$\specline{ [\rtv](\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,m) \ast \mathsf{ContextAt}(old,[\mathsf{rbxv,\ldots,\rtv}]) \ast \mathsf{ContextAt}(new,[\mathsf{rbxv}',\ldots,\rtv']))  }_{\rtv}$
$\specline{ \mathcal{I}\texttt{ASpace}(\theta',m') \ast \texttt{Pother} \ast  \texttt{rsi}\mapsto_{\textsf{r}} new \ast \texttt{rdi}\mapsto_{\textsf{r}} old \ast \ldots }_{\rtv'}$
\end{lstlisting}
\vspace{-1em}
\caption{Basic task switch code that switches address spaces.}
\label{fig:swtch}
\end{figure}
